Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-04-14 text/back_end_KS.qmd
rq2 ses results

@@

@@

For instance, @fig-chicago-rq2-lowSAT shows the contributions of each Geomarket to the pool of purchased prospect profiles from the two orders targeting lower SAT score ranges (1020-1150) by race/ethnicity. The top panel presents the full distribution of the 15,981 total prospects included in these orders by Geomarkets: 11% reside in Chain of Lakes, 10% in Northwest Suburbs, 2% in North Shore, 4% in Evanston and Skokie, 20% in City of Chicago, 25% in Western Suburbs, and 26% in South and Southwest Suburbs. These overall distributions are compared to those in the lower panels of Figure X, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.

For instance, nearly one half (n=7,728) of all prospects whose contact information was purchased in these lower SAT score orders identified as White. With the exception of the City of Chicago, @fig-chicago-rq2-lowSAT shows all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects. The City of Chicago Geomarket, however, contributed only 5% of White prospects, despite accounting for 20% of all included prospects. Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, exept for Evanston and Skokie (12% versus 4% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (10% versus 26% for all prospects) contributing smaller shares relative to their overall representation.

In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT. Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares - 48% and 34%, respectively- compared to their overall contributions of 20% and 26% of all prospects whereas all other Geomarkets contributed substantially smaller proportions. The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions than their overall representation among all included prospects.

@@

```{r fig-chicago-rq2-lowSAT, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, low SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

>>>>

### Geomarket Changes Over Time \[Need better RQ1 heading\]

@@

### Composition of Students Lists by Geomarkets \[Need Better RQ2 Heading\]

@@

For instance, @fig-chicago-rq2-lowSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting lower SAT score ranges (1020-1150) by race/ethnicity. The top panel presents the full distribution of the 15,981 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 11% reside in Chain of Lakes, 10% in Northwest Suburbs, 2% in North Shore, 4% in Evanston and Skokie, 20% in City of Chicago, 25% in Western Suburbs, and 26% in South and Southwest Suburbs. These overall distributions are compared to those in the lower panels of @fig-chicago-rq2-lowSAT-race, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.

Nearly one half (n=7,728) of all prospects whose contact information was purchased in these lower SAT score orders identified as White. With the exception of the City of Chicago, @fig-chicago-rq2-lowSAT-race shows all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects. The City of Chicago Geomarket, however, contributed only 5% of White prospects, despite accounting for 20% of all included prospects. Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, exept for Evanston and Skokie (12% versus 4% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (10% versus 26% for all prospects) contributing smaller shares relative to their overall representation.

In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT-race. Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares - 48% and 34%, respectively- compared to their overall contributions of 20% and 26% of all prospects whereas all other Geomarkets contributed substantially smaller proportions. The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions than their overall representation among all included prospects.

@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores. Both figures demonstrate how representational Geomarket patterns remain relatively consistent SAT scores increase. For instance, @fig-chicago-rq2-midSAT-race shows that the Chain of Lakes, North Shore, and Western Suburbs Geomarkets contributed similar proportions of White and Asian students, yet smaller proportions of Black and Hispanic students, relative to their overall representation among all included prospects. The Northwest Suburbs and Evanston and Skokie similarly contributed a disproportionately higher share of Asian prospects, but lower shares of Black and Hispanic students. Conversely, the South and Southwest Suburbs Geomarket contributed a substantially larger proportion of Black students, while the City of Chicago yielded larger proportions of both Black and Hispanic students- again, relative to their overall representation among all included prospects targeted at middle range SAT scores. Despite declines in the overall number of Black (n=240) and Hispanic (n=950) prospects included in orders filtering for high-ranges of SAT scores (1310-1600), @fig-chicago-rq2-highSAT-race shows overall Geomarket contribution patterns persist.

We also examine the contributions of each Geomarket to the pool of prospects based on first-generation college student status. Similar to race/ethnicity figures above, @fig-chicago-rq2-lowSAT-ses presents the Geomarket distribution for the 16,060 total prospects (whose first-generations status is known) included in low-range SAT score orders (1020-1150). Among the 3,479 prospects whose parents did not attend college, the City of Chicago stands out as the only Geomarket that contributing a disproportionately larger share of first-generation students - 34% in comparison to 20% for all prospects in this SAT range. In contrast, all other Geomarkets contributed disproportionately smaller shares, ranging from a one-percentage-point difference (Chain of Lakes, Northwest Suburbs, North Shore) to a five-percentage-point difference (Western Suburbs). For prospects whose parents attended some college but did not complete their degrees, @fig-chicago-rq2-lowSAT-ses indicates both City of Chicago and South and Southwest Suburbs Geomarkets contribute a disproportionately larger share of first-generation students. These representational patterns are reversed among prospects that are not first-generationcollege students. While all other Geomarkets contributed either comparable (Northwest Suburbs, South and Southwest Suburbs) or disproportionately larger shares (Chain of Lakes, North Sore, Evanston and Skokie, Western Suburbs), the City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (13%) relative to all prospects (20%) in this SAT range.

@fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status become increasingly pronounced in orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores, respectively. At the middle SAT score range, the City of Chicago Geomarket contributes a substantially larger share of first-generation students whose parents did not attend college—38%, compared to 17% of all prospects in this range, representing a 21-percentage-point disparity. This disproportionality increases to a 25-percentage-point difference at the highest SAT score thresholds, where the City of Chicago accounts for 40% of such first-generation students, despite representing only 15% of all prospects in this score range.

@@

```{r fig-chicago-rq2-lowSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Low-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_race.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-chicago-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_race.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-chicago-rq2-highSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, High-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_race.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-chicago-rq2-lowSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Low-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_ses.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-chicago-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_ses.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-chicago-rq2-highSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, High-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_ses.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```
